Multi-modality Imagery Database for Plant Phenotyping

### 2015-05-01 datapaper_calibration.tex
Updates to datapaper_calibration.tex and datapaper_hardware.tex, as well as 2 figures.

The pixel depth measurements have significant noise.  To characterize the noise variance at a particular depth we collected 300 images of a flat surface facing the camera at this depth and calculated the standard deviation at each pixel.  We repeated this for increments of 10cm between 30cm and 90cm.  At each depth we found that the pixel standard deviations were roughly constant at concentric circles around the optical center.  Hence we can summarize the standard deviation as a function of pixel depth and pixel-radius around the optical center, as shown in Fig.~\ref{fig:Noise}.  This captures a large fraction of the depth error for pixels on plant leafs, although additional sources of error include individual pixel biases, object specularities, strong variations in object albedo, mixed pixels on object edges, and very-low signal reflection.

In the recorded $3$D data we averaged five depth images for each record.  Ideally this leads to reducing the standard deviation by $1/\sqrt{5}$, although the actual benefit is less as some sources of error are not reduced by averaging.

Now we noticed that the chamber light shades blocked some of the depth camera field of view, and in doing so reflected some of the IR illumination.  This resulted in an additional bias shift which we measured and removed from the depth data. % FIXME

>>>>

The time of flight depth measurements can have significant noise, and it is useful to both model it and quantify it.  Doing so can lead to strategies to reduce noise as well as providing guidance to algorithms that use the depth measurements.  Our goal in this section is to provide a simple noise model that can predict the empirically observed depth noise on smooth, Lambertian surfaces, such as plant leafs.

The depth noise $\epsilon$ is modeled as the sum of two components, $\epsilon_I$ and $\epsilon_S$: 
\begin{equation}
	\epsilon = \epsilon_I + \epsilon_S\label{eq:epsilon}
\end{equation}
The term $\epsilon_I$ captures the noise component that varies between images taken from a fixed pose of a static scene. On the other hand the term $\epsilon_S$ models noise that is constant between images but varies between pixels. The variance of $\epsilon_I$ can be estimated by observing pixels of a fixed scene over multiple images.  In our experiments we observed a flat, uniform albedo surface perpendicular to the camera at a sequence of depths, and for each depth acquiring 300 images.  Object depth has a large impact on $\epsilon_I$ likely due to reduced reflected signal.  For constant depth, we observed that the primary factor affecting the variance is the pixel radius from the image center.  Physically we expect this dependence is due to a circularly symmetric illuminating beam closely aligned with the optical axis.  Based on these observations $\sigma_I^2$, the variance of $\epsilon_I$, is empirically estimated as a function of object depth and pixel radius as shown in Fig.~\ref{fig:Noise}.

In addition to image-varying noise, there are pixel depth offsets that are constant for images of the same scene so cannot be removed by averaging over many frames.  We model these with $\epsilon_S$.  To estimate its variance, $\sigma_S^2$, we first average over many depth measurement images, in our case 300, to obtain pixel depth estimates that approximately eliminate the effect of $\epsilon_I$.  Then by calculating true ray depths on a known surface, in our case the observed plane, and assuming that $\epsilon_S$ has the same distribution over all pixels, $\sigma_S^2$ is obtained as the variance of the error between averaged depths and known depths.  In our experiments we obtained $\sigma_S=6.5mm$, and found that it was insensitive to changes in depth.

Assuming independence of $\epsilon_I$ and $\epsilon_S$, the expected variance of depth measurements in the data collection can be obtained as follows.  When $N$ images are averaged to obtain a depth image, where $N=5$ for our data collection, the depth variance is predicted to be:
\begin{equation}
	\sigma^2 = \frac{\sigma_I^2}{N} + \sigma_S^2.\label{eq:sigma}
\end{equation}

There are additional sources of noise not captured by this including object specularities, sharp variations in object albedo, mixed-depth pixels on object edges, and cases of very-low signal reflection.  In addition we noticed that the chamber light shades blocked some of the depth camera field of view, and in doing so reflected some of the IR illumination.  This resulted in a small constant depth shift for the pixels.  We measured this shift for each chamber experiment and provide it as an optional correction to the depth images.

### 2015-05-01 datapaper_hardware.tex
Updates to datapaper_calibration.tex and datapaper_hardware.tex, as well as 2 figures.

% For one-column wide figures use
\begin{figure}
  \includegraphics[width=\linewidth,trim=50 230 45 300,clip]{Figures/CameraConfiguration}
\caption{A plot of the three cameras showing their relative configuration and fields of view as obtained through calibration.  Units are in $mm$.  The optical center of the color camera (on the left) defines the world coordinate system.  Close to it is the depth camera.  Its points are projected into the world coordinate system.  On the right is the combined fluorescent and IR camera.}
\label{fig:CameraConfiguration}
\end{figure}

@@

There are a number of limitations to the depth sensor that become the sources of depth errors. 
The primary measurement limitation on the range-to-target is the strength of the reflected beam. 
As a result, dark, matt surfaces are measured reliably only at a close range on the order of $20$ or $30~cm$.  
Highly reflective surfaces also pose problems with direct reflections leading to saturation and highly unreliable depths.  
In addition reflective surfaces at grazing angles are less reliably measured since little signal is reflected. 
Hence in our data portions of the chamber floor visible in Figure~\ref{fig:CameraConfiguration} are highly reflective and have incorrect depths.  %FIXME
Fortunately the primary goal of the depth measurements are to obtain leaf depths, and plants provide good, roughly Lambertian reflections of IR~\cite{Chelle2006219}.  
Therefore for these reasons the non-leaf depth pixels in the $3$D depth data are unreliable and should be ignored in data analysis.  
Another limitation is that the IR illuminator has a slight offset to the left of the sensor, which results in shadows to the right of some objects, as well as mixed pixels on depth discontinuities.  
Both of these can be readily detected as large standard deviations in the depth image.

The imagery data were collected once every hour. %taken within 10 minutes of each other.  %
These include the fluorescent image, the IR reflectance image with the same camera, the color image, the $3$D depth and a confidence image.  
The $3$D depth image is built from the depth sensor by transforming the points into the world coordinates and is expressed in the unit of $mm$.  
The confidence image is the standard deviation of the depth pixels.  
This is useful for identifying pixels at depth discontinuities that are unreliably detected and result in large standard deviations.  
In addition pixels with no response or saturated pixels are assigned with the maximum standard deviation, and should be filtered out.

\begin{figure}
\centering
  \includegraphics[height=5.9cm,trim=110 250 60 260,clip]{Figures/SigmaRadius} \\
\caption{Noise analysis for a depth camera obtained by imaging a flat surface at various depths.  We found that the standard deviation of the pixel depth measurements had a large dependency on both the target depth and the pixel distance from the image optical center.  A radius of $200$ pixels corresponds to the corners of the depth camera, which can be seen to have far larger noise than at the image center for a given depth. }
\label{fig:Noise}
\end{figure}

>>>>

% For one-column wide figures use
\begin{figure}
\begin{tabular}{c}
  \includegraphics[width=\linewidth,trim=50 200 45 250,clip]{Figures/CameraConfiguration} \\
  ($a$)\\
  \includegraphics[width=6cm,trim=100 200 100 200,clip]{Figures/DepthOnBeanRGB} \\
  ($b$)\\
\end{tabular}
\caption{($a$)A plot of the three cameras showing their relative configuration and fields of view as obtained through calibration.  Units are in $mm$, apart from the size of the images which are scaled and plotted at equal resolution.  The optical center of the color camera (left-most) defines the world coordinate system.  Close to it is the depth camera having much lower resolution.  The right-most camera is the combined fluorescent and IR camera. ($b$) $3$D points from the depth camera are projected along their rays into the world coordinates and then projected into the color and fluorescent camera images.  This shows the projection into the color image.  Only points in a rectangular region around the plant in the depth image are selected, and these are further filtered by eliminating points with high standard deviation.  An algorithm requiring only $3$D points on the plant could select only those that fall on the leaf pixels.}
\label{fig:CameraConfiguration}
\end{figure}

@@

There are a number of limitations to the depth sensor that become the sources of depth errors. 
The primary measurement limitation on the range-to-target is the strength of the reflected beam. 
As a result, dark, matt surfaces are measured reliably only at a close range on the order of $20$ or $30~cm$.  
Highly reflective surfaces also pose problems with direct reflections leading to saturation and highly unreliable depths.  
In addition reflective surfaces at grazing angles are less reliably measured since little signal is reflected. 
Fortunately the primary goal of the depth measurements are to obtain leaf depths, and plants provide good, roughly Lambertian, reflections of IR~\cite{Chelle2006219}.  Thus the depth pixels that are most reliable and are of most use are those that fall on plant foliage.

The imagery data were collected once every hour.
These include the fluorescent image, the IR reflectance image with the same camera, the color image, the $3$D depth and a confidence image.  
The $3$D depth image is built from the depth sensor by transforming the points into the world coordinates and is expressed in the unit of $mm$.  
The confidence image is the standard deviation of the depth pixels.  
This is useful for identifying pixels at depth discontinuities that are unreliably detected and result in large standard deviations.

\begin{figure}
\centering
  \includegraphics[height=5.9cm,trim=110 250 60 260,clip]{Figures/SigmaRadius} \\
\caption{Noise analysis for a depth camera obtained by imaging a flat surface at various depths.  We found that the standard deviation, $\sigma_I$ from Eq. (\ref{eq:sigma}) of the pixel depth measurements had a large dependency on both the target depth and the pixel distance from the image center, and these are plotted.  A radius of $0$ pixels is the image center, and of $200$ pixels corresponds to the corners of the depth camera, which can be seen to have far larger standard deviation than at the image center for the same depth. }
\label{fig:Noise}
\end{figure}
